---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# DICOM 
DICOM (Digital Imaging and Communications in Medicine) format

- Standardized representation of images
- Formats of files from scanner to hospital PACS (picture archiving and communication system)
- Two components: image data (in pixles) and header (metadata)

## readDICOM Output

The pixel data in a DICOM file is a matrix

One DICOM file = one "slice of the brain

The `oro.dicom` package enables reading DICOM files in R. 


Each slice is stored as a list.

```{r}
library(oro.dicom)
slice = readDICOM("Neurohacking_data/BRAINIX/DICOM/FLAIR/IM-0001-0011.dcm")
suppressMessages(class(slice))
```

This contains two lists: `hdr` and `img`

```{r}
names(slice)
class(slice$hdr)
class(slice$img)
```

Each element of `hdr` is a data.frame and each element of `img`  is a matrix

```{r}
class(slice$hdr[[1]])
class(slice$img[[1]])
dim(slice$img[[1]])
```

## DICOM Images

We can look at the image in R.

```{r}
d = dim(t(slice$img[[1]]))
image(1:d[1], 1:d[2], t(slice$img[[1]]), col=gray(0:64/64))
```

For analysis we want to look at the numbers in the matrix, not the image in the grayscale.

### Working with the numbers

```{r}
slice$img[[1]][101:105, 121:125]
```

And we can plot the data as usual using histograms
```{r}
hist(slice$img[[1]][,], #all elements
breaks=50, xlab="FLAIR",
prob=T, col = rgb(0, 0, 1, 1/4),
main="")
```

## DICOM Header Information
Now let's look at the header.

```{r}
hdr = slice$hdr[[1]]
str(hdr)
```

Each row of the dataframe has a name:

```{r}
hdr$name
```

Let's look at the `"PixelSpacing"` for this images, which represents the resolution of the image. 

```{r}
hdr[hdr$name == 'PixelSpacing', 'value']
```

We can also look at the Flip Angle used while collecting the image.
```{r}
hdr[hdr$name == 'FlipAngle', ]
```

## Loading Multiple DICOM Files

For analysis, we will need to read in all the slices. Specifying a directory instead of a file name in `readDICOM` will load all images in that directory as a single object.
```{r}
all_slices_T1 = readDICOM("Neurohacking_data/BRAINIX/DICOM/T1/")
```

Examining the dimensions of the first slice shows that the T1 image has a higher resolution than the FLAIR image. Thus this is a bigger
```{r}
dim(all_slices_T1$img[[1]])
hdr = all_slices_T1$hdr[[1]]
hdr[hdr$name == "PixelSpacing", "value"]
```

## Summary 
-DICOM files contain a lot of information
-Image data are stored as collection of 2D slice files
-Different sequences can have different resolutions
-Data and metadata can be accessed in R
-Contains protected health data
-Structure is cumbersome for 3D analysis
-A genuine 3D format is needed:: NIfTI














