---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NIfTI

NifTI (Neuroimaging Informatics Tehcnology Image)
-Standardized representation of images
-Most commonly used tpy eof analytic file
-Developed to facilitate cross-platform, cross-software interpretability
-3-dimensional array: stacking individual slices on top of each other
-DICOM: one sheet of paper; NIfTI: stack of papers

## From DICOM to NIfTI

-DICOM to NIFTI using the `dicom2nifti` function in the `oro.dicom` package
-The `nifti` object becomes an R object
-After saving a file to a `nifti` file it can be used without R

Let's read all DICOM images:

```{r}
library(oro.dicom)
all_slices_T1 = readDICOM("Neurohacking_data/BRAINIX/DICOM/T1")
dim(all_slices_T1$img[[11]])
hdr = all_slices_T1$hdr[[11]]
hdr[hdr$name == "PixelSpacing", "value"]


```

Now lets's convert them to NIFTI:
```{r}
nii_T1 = dicom2nifti(all_slices_T1)
d = dim(nii_T1); d; class(nii_T1)
``` 

And plot one slice:
```{r}
# Plot the 11th slice
image(1:d[1], 1:d[2], nii_T1[,,11], col=gray(0:64/64), xlab="", ylab="")
```

## Write and Read NIfTI files

- Use the `writeNIfTI, readNIfTI` functions in the `oro.nifti` package
- Default for `writeNIfTI` is to save the compressed NIfTI files

```{r}
library(oro.nifti)
dir_name = "Neurohacking_data/BRAINIX/NIfTI/"
fname = "Output_3D_File"
writeNIfTI(nim=nii_T1, file = paste0(dir_name, fname))
```

And check that the file si there after saving
```{r}
list.files(dir_name, pattern = fname)
```

Now lets read a different file:

```{r}
list.files(dir_name, pattern="T") #Check that file is in directory
nii_T2=readNIfTI(paste0(dir_name,"T2.nii.gz"), reorient=FALSE)
dim(nii_T2)
```

## Compressed Image Files

-A non-compressed file can be obtained using the argument `gzipped=FALSE` in the function `writeNIfTI`






