---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Visualization

This lecture goes over basic visualization methods. Objectives are the following:  
-Visualize Slices
-Digital Zoom
-Histograms of Intensities
-Back mapping

## Read NIfTI

```{r}
library(oro.nifti)
dir = "Neurohacking_data/BRAINIX/NIfTI/"
fname = "Output_3D_File"
fpath = paste0(dir,fname)
(nii_T1 = readNIfTI(fname = fpath))
```

Printing the NIfTI object show the structure of the data

## Visualizing Slices

As a first pass, we could use the image function from the graphics package.

```{r}
#Save dimensions of the image
d = dim(nii_T1)

# Visualizing the 11th axial slice
graphics::image(1:d[1], 1:d[2], nii_T1[,,11], 
      col = gray(0:64/64) xlab="", ylab="") 
```

However, `oro.nifti` has its own image function which is called by default one NIfTI objects.

```{r}
image(nii_T1, z = 11, plot.type = "single")
```

If no arguments are passed for coordinates, `oro.nifti::image` plots all the slices axially. 

```{r}
image(nii_T1)
```

The `orthographic` functions allows to visualize a single point from three dimensions. 

```{r}
orthographic(nii_T1, xyz = c(200,220,11))
```

## Exploratory histogram of intensities

Using R base graphics, we can get a quick view of the intensites in the brain.
```{r}
par(mfrow = c(1, 2));
o = par(mar = c(4,4,0,0))
hist(nii_T1, breaks = 75, prob = T, xlab="T1 intensities", col = rgb(0,0,1, 1/2), main="")

# Plot filtered by > 20
hist(nii_T1[nii_T1 > 20], breaks = 75, prob = T, xlab="T1 intensities > 20", col = rgb(0,0,1, 1/2), main="" )
```

A lot ot the voxels, are ~0 in the first plot since most of the image is the black background. The second plot shows the intensities that are higher than 20.

## Back mapping

Backmapping consists of overlaying a mask over an image of the brain. A basic backmapping can be achieved by taking creating a mask according to intensities and calling the function `overlay`.

```{r}
is_btw_300_400 = nii_T1 > 300 & nii_T1 < 400
nii_T1_mask = nii_T1
nii_T1_mask[!is_qbtw_300_400] = NA
overlay(nii_T1, nii_T1_mask, z = 11, plot.type="single")
```

`overlay` can be called to show a single slice as above, or the whole brain as below.

```{r}
overlay(nii_T1, nii_T1_mask)
```


```{r}
orthographic(nii_T1, nii_T1_mask, xyz=c(200,220,11), text="Image overlaid with a mask", text.cex = 1.5)
```






 