---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Manipulation

This lecture covers:
- Masking
- Operations

## Download data from GitHub repository

The code below downloads (1) the brain data and (2) the mask for this data.
```{r}
"This code chunk has been adapted from from https://github.com/muschellij2/Neurohacking/blob/master/week2/Basic%20Data%20Manipulation.Rmd"

url = "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-01-MPRAGE.nii.gz"
fname = "SUBJ0001-01-MPRAGE.nii.gz"
fpath = file.path("Neurohacking_data/kirby21", fname)
download.file(url, fpath, mode="wb")

maskurl = "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001_mask.nii.gz"
maskfname = "SUBJ0001_mask.nii.gz"
maskfpath = file.path(getwd(), maskfname)
download.file(maskurl, maskfpath, mode="wb")# NIfTI is binaryfile format

library(oro.nifti)
T1 = readNIfTI(fpath,reorient=FALSE)
mask = readNIfTI(maskfpath, reorient=FALSE) 
```

## Visualization

This brain was collected with a bigger FOV than the previous scans. Thus a lot of body parts other than brain are shown. 
```{r}
orthographic(T1)
```


```{r}
orthographic(mask)
```

## Masking

```{r}
marked.T1 = T1*mask
orthographic(masked.T1)
```

##  Substraction

Load the follow up T1 scan from the Girthub repository

```{r}
url = "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-02-MPRAGE.nii.gz"
fname = "SUBJ0001-02-MPRAGE.nii.gz"
fpath = file.path("Neurohacking_data/kirby21", fname)
download.file(url, fpath, mode="wb")

T1.follow  = readNIfTI(fpath, reorient = FALSE)

substract.T1 = T1.follow - T1
min(substract.T1)
max(substract.T1)
```

```{r}
orthographic(substract.T1)
```

