---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r read}
#set fsl path before running fslr in RStudio
library(fslr)
if(!have.fsl()){
  options(fsl.path = '/usr/local/bin')
} 

#Check file paths for data downloaded from
library(oro.nifti)
fpath = "Neurohacking_data/kirby21/visit_1/113"
fname = "113-01-MPRAGE"
nim = readNIfTI(file.path(fpath,fname), reorient=FALSE)
```

#Some statistics using FSL

```{r fslstats}
mean(nim)

# can call fsl within r on the 'nim' img
fslstats(nim, opts="-m")

# can also provide a filepath instead of r object
fslstats(fpath, opts="-m")
fslstats.help()
```
# Perform inhomogeneity correction

```{r fast}
# fsl_biascorrect calls FAST from FSL
fast_img = fslr::fsl_biascorrect(nim, retimg=TRUE)
```

#Plot difference between images in R

we can use colors to visually inspect the difference between two images. In this case, the bias-corrected image was substracted from the original image and the resulting intensities (i.e. areare colored based on their sign. Quantiles are used to provide to derive the color palette. 

```{r plotdiff}
sub_bias = niftiarr(nim, nim-fast_img)

#quantile the difference image using these breaks
q = quantile(sub_bias[sub_bias != 0], probs = seq(0,1,0.1))

#install.packages("scales")
library(scales)

#get a gradient palette to map onto divergence between images
fcol = scales::div_gradient_pal(low="blue", mid="yellow", high="red")

ortho2(nim, sub_bias,
       col.y=alpha(fcol(seq(0,1, length = 10))), #color for second arg
       ybreaks= q,
       ycolorbar= TRUE,
       text = paste0("Original  Image Minus N4", "\n Bias-Corrected Image")
)
```

# Plotting histograms of intensities for original and bias-corrected images

```{r plot-intensity}

# choose a subset of slices to be plotted
slices = c(2,6,10,14,18)

# for each slice number save that slice from the original and bias-corrected images into an array
vals = lapply(slices, function(x){
  cbind(img = c(nim[,,x]),
        fast = c(fast_img[,,x]),
        slice = x)
})

# row bind all elements of the vals list to get a data.frame
vals = do.call("rbind", vals)
vals = data.frame(vals)

# filter out values below 0
vals = vals[vals$img > 0 & vals$fast > 0,]

# Change column names
colnames(vals)[1:2] = c("OriginalValue", "Bias-Corrected Value")
v = reshape::melt(vals, id.vars = "slice")

library(dplyr)
library(ggplot2)
g = v %>% ggplot(aes(x = value,
                     col = factor(slice))) +
  geom_line(stat = "density") +
  facet_wrap(~variable) +
  scale_colour_discrete(name = "Slice #")

g
```

After bias field correction the intensity curves line up better across slice. Also, the relative intensities are affected such that the highest intensity. 

# Write NIfTi

```{r write}
fname = paste0("fast-",fname)
suppressMessages(oro.nifti::writeNIfTI(fast_img, file.path(fpath, fname)))
```

