---
output: html_document
---

# Skull Stripping with FSL's Brain Extraction Tool (BET)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading file

```{r}
#set fsl path before running fslr in RStudio
library(fslr)
if(!have.fsl()){
  options(fsl.path = '/usr/local/bin')
} 

#Check file paths for data downloaded from
library(oro.nifti)
fpath = "Neurohacking_data/kirby21/visit_1/113/"
fname = "fast-113-01-MPRAGE.nii.gz"
fast_img = readNIfTI(paste0(fpath,fname), reorient=FALSE)
```

## Running BET

```{r}
bet_fast = fslr::fslbet(infile=fast_img, retimg = TRUE)
```

## Visualizing first past

To visualize the effects of the brain extraction on the original, the `orthographic` may be used twice by passing (1) the image returned by BET, (2) the original image overlaid by a mask. To produce the mask, follow the code below. 
```{r}
# create an array of 1s with same dimensions as the skull-stripped image
bet_fast_mask = niftiarr(bet_fast, 1)

# set all non-brain voxel to NA
bet_fast_mask[!(bet_fast > 0)] <- NA

# plot
orthographic(bet_fast)
orthographic(fast_img, bet_fast_mask)
```

The results are not great since a lot of non-brain areas get included.

## Improving Brain Extraction

To improve upon the results, a second pass can be done taking into account the center of gravity. 

```{r}
# use the cog function on the skull-stripped image
cog = cog(bet_fast, ceil = TRUE)

#prepare a string to be passed to opts parameter of fslbet
cog = paste("-c", paste(cog, collapse = " "))

# run BET on the bias-corrected img using the center of gravity to get cleaner results
bet_fast2 = fslbet(infile = fast_img, retimg = TRUE, opts = cog)
```

## Visualizing first and second pass
```{r}
orthographic(bet_fast)
orthographic(bet_fast2)
```

The results have improved, removing much of the lower non-brain sections. However, we can see the results are not perfect so it's important to visually inspect each image at each phase of the processing pipeline.
